---
title: "Game and Grade"
subtitle: "Do Video Games Impact Academic Performance"
author: "Carolyn Powers"
institute: "Houghton University"
date: "5/15/25"
output:
  xaringan::moon_reader:
    css: custom-theme.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(dplyr)
library(ggplot2)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
#| label: sourcedata

# From Kaggle: https://www.kaggle.com/datasets/deadier/play-games-and-success-in-students
gameandgrade <- read_csv(file = "data/gameandgrade.csv")
gameandgrade
```

```{r, include=FALSE}
#Xaringan image for slides 
style_xaringan(
  title_slide_background_image = "img/controllerBackground.jpg",
  background_image = "img/controllerBackground.jpg"
)
```

<span style="color: white;">
Sex: The student's gender ('Female' or 'Male').
<span style="color: white;">

School Code: The code representing the student's school, which is a unique identifier for each school.

Playing Years: The number of years the student has been playing games.

Playing Often: A rating of how often the student plays games, likely on a scale from 1 to 5.

Playing Hours: The number of hours the student spends playing games per day.

Playing Games: The number of types of games the student plays (e.g., 0: not playing games, 1: playing games).

Parent Revenue: The annual income of the student's family.

Father Education: The education level of the student's father.

Mother Education: The education level of the student's mother.

Grade: The student's academic grade.
</span>
---

class: top, left

#Limitations

--

##-Frequency 
##-"Good Games"

---

```{r, echo=FALSE, include=FALSE}
#looked up how to change a column to numbers, it had NAs, so looked up how to remove them.
  gameandgrade$Grade <- gsub("[^0-9.]", "", gameandgrade$Grade)
  gameandgrade$Grade <- as.numeric(gameandgrade$Grade)
```

```{r, echo=FALSE, include=FALSE}
#the average score of the schools we are looking at, all are schools that if you round up are 70 or above.
gameOrNotGameGrades <- gameandgrade %>%
  filter(`Playing Games` != "2") %>%
  group_by(`Playing Games`) %>%
  summarise(`Average Score` = mean(Grade, na.rm = TRUE))
```

#Number of schools and their Grades

```{r, echo=FALSE}
#Helpful code to make slides look better
knitr::kable(gameOrNotGameGrades, align = 'l')
```

---

```{r, echo=FALSE, include=FALSE}
#Average score of schools based on school code/summarizing, adding mean
averageScoreofSchools <- gameandgrade %>%
  filter(`Playing Games` == "1") %>%
  group_by(`School Code`) %>%
  summarize(`Average Score` = mean(Grade, na.rm = TRUE))
```

#Number of schools and their Grades

```{r, echo=FALSE}
#Adding slide coding, the average grade of students who play video games in their respective schools
knitr::kable(averageScoreofSchools, align = 'l')
```

---

```{r, echo=FALSE}
#Playing hours special characters 
  gameandgrade$`Playing Hours` <- gsub("[^0-9.]", "", gameandgrade$`Playing Hours`)
  gameandgrade$`Playing Hours` <- as.numeric(gameandgrade$`Playing Hours`)
```

```{r, echo=FALSE, include=FALSE}
#Seeing playing hours variable/finding acerage score
gameandgrade %>%
  group_by(`Playing Hours`) %>%
  summarise(averageScore = mean(`Grade`, na.rm = TRUE))
```

```{r, echo=FALSE}
#Vizualizing data found
gameandgrade %>%
  group_by(`Playing Hours`) %>%
  summarise(averageScore = mean(`Grade`, na.rm = TRUE)) %>%
  ggplot(aes(x = `Playing Hours`, y = averageScore, fill = `Playing Hours`)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Score based on hours played per day", x = "Hours Played", y = "Average Grades") +
  theme_minimal()
```

---

```{r, echo=FALSE}
#Boys play more on average
gameandgrade %>%
  group_by(`Sex`) %>%
  summarise(averageScore = mean(`Playing Hours`, na.rm = TRUE))
```

```{r, echo=FALSE}
#Girls on average have a higher Grade score
gameandgrade %>%
  group_by(`Sex`) %>%
  summarise(averageScore = mean(`Grade`, na.rm = TRUE))
```

---

```{r, echo=FALSE}
#Converting parents education from a string
  gameandgrade$`Father Education` <- gsub("[^0-9.]", "", gameandgrade$`Father Education`)
  gameandgrade$`Father Education` <- as.numeric(gameandgrade$`Father Education`)
  gameandgrade$`Mother Education` <- gsub("[^0-9.]", "", gameandgrade$`Mother Education`)
  gameandgrade$`Mother Education` <- as.numeric(gameandgrade$`Mother Education`)
```

```{r, echo=FALSE}
#Combining parent education 
gameandgrade <- gameandgrade %>%
mutate(combinedParentEduc = `Father Education` + `Mother Education`)
```

```{r, echo=FALSE}
#Violin Chart
gameandgrade %>%
  group_by(combinedParentEduc) %>%
  summarise(averageScore = mean(`Grade`, na.rm = TRUE)) %>%
  ggplot(aes(x = combinedParentEduc, y = averageScore)) +
  geom_violin(fill = "skyblue", color = "black") +
  labs(title = "1", x = "2", y = "3")
```

---

```{r, echo=FALSE}
#Bar Graph2
gameandgrade %>%
  filter(combinedParentEduc != 0 & combinedParentEduc != 1) %>%
  group_by(combinedParentEduc) %>%
  summarise(averageScore = mean(`Grade`, na.rm = TRUE)) %>%
  ggplot(aes(x = combinedParentEduc, y = averageScore, fill = combinedParentEduc)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Score based on hours played per day", x = "Combined Parent Education Score", y = "Average Grades") +
  theme_minimal()
```
 
---

```{r, echo=FALSE}
#Bar Graph no2
gameandgrade %>%
    filter(combinedParentEduc != 0 & combinedParentEduc != 1 & combinedParentEduc != 2) %>%
  group_by(combinedParentEduc) %>%
  summarise(averageScore = mean(`Playing Hours`, na.rm = TRUE)) %>%
  ggplot(aes(x = combinedParentEduc, y = averageScore, fill = combinedParentEduc)) +
  geom_bar(stat = "identity") +
  labs(title = "Do more educated parents allow their children to play less?", x = "Combined Parent Education Score", y = "Average Daily Hours") +
  theme_minimal()
```

```{r, echo=FALSE, include=FALSE}
#Parent revenue vs. average scores
gameOrNotGameGrades <- gameandgrade %>%
  filter(`Parent Revenue` != "2") %>%
  group_by(`Parent Revenue`) %>%
  summarise(`Average Score` = mean(Grade, na.rm = TRUE))
```

---

```{r, echo=FALSE}
#Vizualizing data found above
gameandgrade %>%
  group_by(`Parent Revenue`) %>%
  summarise(averageScore = mean(`Grade`, na.rm = TRUE)) %>%
  ggplot(aes(x = `Parent Revenue`, y = averageScore, fill = `Parent Revenue`)) +
  geom_bar(stat = "identity") +
  labs(title = "Parent Revenue influence on average score", x = "Parent Revenue", y = "Average Grades") +
  theme_minimal()
```

---

Other stuff I need:

Add own images 

fix slides (knitr::kable) makes them pretty 

game and grade average score of schools 

make more slides 

fix outlier lines 138-147


Works Cited
Games and success in students. (2025, February 19). Kaggle. https://www.kaggle.com/datasets/deadier/play-games-and-success-in-students